---
title: "Homework1 Sam Whitttaker"
author: "Sam Whittaker"
date: "1/26/2022"
output: html_document
---
Goal: To test the hypothesis of the paper: "Hibernation constrains brain size evolution in mammals" through 

2nd Goal: To test my own hypothesis that brain size is affected by their diet type
Data shown below 
```{r}
head(Supplementary_material)
summary(Supplementary_material)
```
First place to investigate is to see if any correlation can be determined through a box plot. 

```{r, fig.height=2.5}
`Hibernation (0=No, 1=Yes)`<-as.character(Supplementary_material$hibernation)
`Brain Mass (g)`<- Supplementary_material$`brain mass`
ggplot(Supplementary_material, aes(x = `Hibernation (0=No, 1=Yes)`, y = `Brain Mass (g)`)) + 
   geom_boxplot() + coord_flip()
```
*Borrowed akraf's idea to use coord_flip() to make the box plot look nicer* 

While there are many more sample from the non-hibernating animals and a wider range of values, a correlation is still visable. The boxplot does show that the non-hibernating animals have a larger brain mass even without including the 2 outlires. This indicates that hibernation may in fact limit brain mass. 

A statistical test needs to be done in order to further show that this hypothesis is supported. The test of choice is a Welch Two Sample t-test.

```{r}
t.test(Supplementary_material$`brain mass`, Supplementary_material$hibernation)
```

With p-value < 2.2e-16 we can reject the null hypothesis meaning that the effect of hibernation on brain mass is significant. So hibernation resluting in lower brain mass is supported by the data. 

Goal 1: completed 
Goal 2: To test my own hypothesis that brain size is affected by their diet type.

```{r, fig.height=2.5}
Diet<-as.character(Supplementary_material$diet)
ggplot(Supplementary_material, aes(x = Diet, y = `Brain Mass (g)`)) + 
   geom_boxplot() + coord_flip()
```

At first glance a correlation is not very clear due to the large outlines amount herbivores. At further inspection you can see that herbivores and carnivores have slightly larger brain masses then those of omnivores and garnivores. Further signfigance testing is needed to see the relationship. 

```{r}
t.test(Supplementary_material$`brain mass`, Supplementary_material$diet)
```
With p-value < 2.2e-16 we can reject the null hypothesis meaning that the effect of diet on brain mass is significant. The type of diet is shown to be correlated with brain mass. 
