---
title: "Homework1 Sam Whitttaker"
author: "Sam Whittaker"
date: "1/26/2022"
output: html_document
---
Goal: To test the hypothesis of the paper: "Hibernation constrains brain size evolution in mammals" through 

2nd Goal: To test my own hypothesis that brain size is affected by their diet type
Data shown below 
```{r}
head(Supplementary_material)
summary(Supplementary_material)
```
First place to investigate is to see if any correlation can be determined through a box plot. 

```{r, fig.height=2.5}
`Hibernation (0=No, 1=Yes)`<-as.character(Supplementary_material$hibernation)
`Brain Mass (g)`<- Supplementary_material$`brain mass`
ggplot(Supplementary_material, aes(x = `Hibernation (0=No, 1=Yes)`, y = log10(`Brain Mass (g)`))) + 
   geom_boxplot() +
   theme_classic() +
  labs(x= "Hibernation", y= "log(Brain Mass (g))", title = "Hibernation Affect on Brain Mass", caption = "0 = Non-Hibernating, 1 = Hibernating") 

```

While there are many more sample from the non-hibernating animals and a wider range of values, a correlation is still visable. The boxplot does show that the non-hibernating animals have a larger brain mass even without including the 2 outlires. This indicates that hibernation may in fact limit brain mass. 

A statistical test needs to be done in order to further show that this hypothesis is supported. The test of choice is a Welch Two Sample t-test.

```{r}
t.test(Supplementary_material$`brain mass`, Supplementary_material$hibernation)
```

With p-value < 2.2e-16 we can reject the null hypothesis meaning that the effect of hibernation on brain mass is significant. So hibernation resulting in lower brain mass is supported by the data. The data is likely skewed however due to two brain masses that are over 10x larger then most of the other brain mass measurements. 

Goal 1: completed 
Goal 2: To test my own hypothesis that brain size is affected by their diet type.

```{r, fig.height=2.5}
Diet<-as.character(Supplementary_material$diet)
ggplot(Supplementary_material, aes(x = Diet, y = log10(`Brain Mass (g)`))) + 
   geom_boxplot() +
  theme_classic() +
  labs(x= "Diet", y= "log(Brain Mass (g))", title = "Diets Affect on Brain Mass", caption = "0 = Herbivore, 1 = Garnivore, 2 = Omnivore, 3 = Carnivore") 
  
```

At first glance a correlation is not very clear due to the large outlines amount herbivores. At further inspection you can see that herbivores and carnivores have slightly larger brain masses then those of omnivores and garnivores. Further significance testing is needed to see the relationship. 

```{r}
`Anova Diet` <- aov(diet ~ `brain mass`, data=Supplementary_material)
summary(`Anova Diet`)
```
With a Pr(>F) of 0.00466 we can reject the null hypothesis meaning that the effect of diet on brain mass is significant. The type of diet is shown to be correlated with brain mass. Data is likely very skewed however, due to two herbivores with brain masses that are over 10x as large as most of the other brain masses. Body mass to brain ratio is not considered in this test and would likely provide a more clear answer on if brain mass is affected by diet type.

*Worked with Olivia Dyer to fix the box plots by adding log10 to them*
*Used PDFs in cheat sheet folder on github page for a lot of the code*